---
title: "ConfidenceIntervals-HypothesisTesting"
author: "Rahul Sinha"
date: 2018-10-24
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

# Confidence Intervals
a. A sample of size n = 100 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 4, compute a 95% confidence interval for the population mean μ.
```{r}
n = 100
meanXbar = 59
se = 4/sqrt(n)
confidence_level = 0.95
# the sample size is 100, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 95% confidence interval
lower_bound
upper_bound

vec_sample_size <- c()
vec_sample_size <- c(vec_sample_size, n)
vec_width <- c()
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


b. A sample of size n = 100 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 4, compute a 90% confidence interval for the population mean μ.
```{r}
n = 100
meanXbar = 59
se = 4/sqrt(n)
confidence_level = 0.90
# the sample size is 100, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 90% confidence interval
lower_bound
upper_bound

vec_sample_size <- c(vec_sample_size, n)
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


c. A sample of size n = 100 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 4, compute a 99% confidence interval for the population mean μ.

```{r}
n = 100
meanXbar = 59
se = 4/sqrt(n)
confidence_level = 0.99
# the sample size is 100, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 99% confidence interval
lower_bound
upper_bound

vec_sample_size <- c(vec_sample_size, n)
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


d. A sample of size n = 25 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 4, compute a 95% confidence interval for the population mean μ.
```{r}
n = 25
meanXbar = 59
se = 4/sqrt(n)
confidence_level = 0.95
# the sample size is 25, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 95% confidence interval
lower_bound
upper_bound

vec_sample_size <- c(vec_sample_size, n)
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


e. A sample of size n = 400 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 4, compute a 95% confidence interval for the population mean μ.
```{r}
n = 400
meanXbar = 59
se = 4/sqrt(n)
confidence_level = 0.95
# the sample size is 400, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 95% confidence interval
lower_bound
upper_bound

vec_sample_size <- c(vec_sample_size, n)
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


f. A sample of size n = 100 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 1, compute a 95% confidence interval for the population mean μ.
```{r}
n = 100
meanXbar = 59
se = 1/sqrt(n)
confidence_level = 0.95
# the sample size is 100, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 95% confidence interval
lower_bound
upper_bound

vec_sample_size <- c(vec_sample_size, n)
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


g. A sample of size n = 100 produced the sample mean of X̄ = 59. Assuming the population standard deviation σ = 16, compute a 95% confidence interval for the population mean μ.
```{r}
n = 100
meanXbar = 59
se = 16/sqrt(n)
confidence_level = 0.95
# the sample size is 100, hence we safely assume a normal sampling distribution of the mean.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = meanXbar + (critical_value * se)
lower_bound = meanXbar - (critical_value * se)

# 95% confidence interval
lower_bound
upper_bound

vec_sample_size <- c(vec_sample_size, n)
width = upper_bound - lower_bound
vec_width <- c(vec_width, width)
```


We see that the width of the confidence interval in part a is ~ 1.56. For part b the confidence level decreases to 90% this results in shorter width, ~ 1.32, as to be less confident of capturing the population mean in our interval we can decrease our confidence interval.
Similarly for part c to be sure of capturing the population mean 99% of the times in our samples, we must increase our confidence interval. Thus the width for part c increases to 2.06. 
For part d, the width increases futher to 3.14; this is due to the reduction in sample size which results in an increased standard error; hence the bigger margin or width of confidence interval.
For part e, compared to a, the sample size quadruples which results in a smaller standard error and hence its confidence interval width decreases to 0.78 compared to 1.57 of part a.
For part f, the standard deviation of the population reduces to 1 which reduces the standard error, hence reducing the width of confidence interval to 0.39.
Part g, the population standard deviation increases to 16 resulting in an increased standard error, hence increasing the width of confidence interval to 6.27.

```{r}
plot(vec_sample_size, vec_width, pch=21, bg = "red")
abline(h=vec_width[1], lty=2, col = "green")
text(vec_sample_size[1],vec_width[1], "a:sd4  95%")
text(vec_sample_size[2],vec_width[2], "b:sd4  90%")
text(vec_sample_size[3],vec_width[3], "c:sd4  99%")
text(vec_sample_size[4],vec_width[4], "d:sd4  95%")
text(vec_sample_size[5],vec_width[5], "e:sd4  95%")
text(vec_sample_size[6],vec_width[6], "f:sd1  95%")
text(vec_sample_size[7],vec_width[7], "g:sd16 95%")
vec_sample_size
vec_width
```


# Confidence Intervals
```{r}
real_estate <- data.frame(read.csv("Data/Real_estate.csv"))
```

```{r}
n = 50
meanPrices = mean(real_estate$prices)
se = sd(real_estate$prices)/sqrt(n)
confidence_level = 0.95
# the sample size is 50, and the prices are normally distributed.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = (meanPrices + (critical_value * se)) * 1000
lower_bound = (meanPrices - (critical_value * se)) * 1000

# 95% confidence interval for the mean value of homes
lower_bound
upper_bound
```

```{r}
n = 50
meanPrices = mean(real_estate$prices)
se = sd(real_estate$prices)/sqrt(n)
confidence_level = 0.92
# the sample size is 50, and the prices are normally distributed.
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = (meanPrices + (critical_value * se)) * 1000
lower_bound = (meanPrices - (critical_value * se)) * 1000

# 92% confidence interval for the mean value of homes
lower_bound
upper_bound
```

```{r}
# n = ?
meanPrices = mean(real_estate$prices)
# se = sd(real_estate$prices)/sqrt(n)
confidence_level = 0.95
width = 100000
# the sample size is 50, and the prices are normally distributed.
critical_value = qnorm(1-(1-confidence_level)/2)
# 2 * critical_value * Standard Error = 100,000
# Standard Error = population standard deviation/sqrt(n)
# standard deviation of the given prices table gives a good estimation of the population standard deviation.
# Standard Error = sd(real_estate$prices)/sqrt(n)

# given: confidence interval width = 100000
# (critical_value * se) * 2 = 100,000
# (critical_value * sd(real_estate$prices) * 1000/sqrt(n)) * 2 = 100,000
# sqrt(n) = (critical_value * sd(real_estate$prices) * 1000 * 2 / 100000)
# n = (critical_value * sd(real_estate$prices) * 1000 * 2 / 100000)^2
n = (critical_value * sd(real_estate$prices) * 1000 * 2 / width)^2
cat("calculated sample size =", n, "\n")
cat("rounded sample size =", ceiling(n))
```

# Hypothesis Testing

```{r}
library(ISwR)
attach(graft.vs.host)
```

```{r}
# patients_gvhd 
X = 17
n = 37

# checking if nπ >= 5 and n(1-π) >= 5
n * X/n >= 5
n * (1 - (X/n)) >= 5
# Thus z-distribution valid for this problem

```

```{r}
# proportion of patients having gvhd in the sample
p_hat = X/n
se = sqrt(p_hat*(1-p_hat)/n)
confidence_level = 0.98
# the sample size is 37, using z statistics
critical_value = qnorm(1-(1-confidence_level)/2)
upper_bound = p_hat + (critical_value * se)
lower_bound = p_hat - (critical_value * se)

# 98% confidence interval for the population proportion of the number of people that are given this bone transplant that end up having the graft-versus-host disease
lower_bound
upper_bound
```

```{r}
# proportion of patients having gvhd in the sample
# Suppose p_hat = X/n
# se = sqrt(p_hat*(1-p_hat)/n)
confidence_level = 0.98
# the sample size is 37, using z statistics
critical_value = qnorm(1-(1-confidence_level)/2)
# We know that Margin of error = critical_value * standard error
# critical_value * standard error < 0.05
# critical_value * sqrt(p_hat*(1-p_hat)/n) < 0.05
# sqrt(p_hat*(1-p_hat)/n) < 0.05/critical_value
# p_hat*(1-p_hat)/n < (0.05/critical_value)^2
# n > p_hat*(1-p_hat)/(0.05/critical_value)^2
# we know that the max value of y*(1-y) = 0.5 * 0.5
n = 0.5*0.5/(0.05/critical_value)^2

# thus sample size n = 542
cat("calculated sample size =", n, "\n")
cat("rounded sample size =", ceiling(n))

```

