---
title: "Probability Distributions"
author: "Rahul Sinha"
date: 2018-09-26
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

# Binomial Distribution
Probability that a student matriculates if he is offered admission = 0.15


Let X be the discrete random variable representing the number of students who matriculate.
Every student offered admission is a trial with the successful event being his enrollment.
The probability of this success for every student = π = 0.15
Hence X is a Binomial Random Variable with n = 1000 and π = 0.15
Number of trials is represented by the number of students offered admission.
Number of students university offers admission to = n = 1000
Expected number of students that will enroll or E[X] = n x π = 1000 x 0.15 = 150


For a binomial distribution of Discrete Random Variable X:
Standard Deviation of X = √[(nπ)(1 - π)] = √(1000 x 0.15 x (1 - 0.15)) = 11.29
```{r}
sqrt(1000 * 0.15 * (1 - 0.15))
```


The university wants to be 99.5% sure,
P(X <= 160) > .995
Through Trial and Error using the pbinom function, we find that 883 offers should be made to be 99.5% sure that no more than 160 students matriculate.
```{r}
pbinom(160, size = 883, prob = 0.15)
```


# Airline Revenue Optimization
Let X measure the number of passengers that show up.
Each passenger shows up with probability of 0.93


n = 73
π = 0.93
Thus X is a Random Variable with n = 73 and π = 0.93
P(X > 70 | n,π) = 1 - P(X<=70) = 1 - pbinom(70, size = 73, prob = 0.93) = 0.1069867
```{r}
1 - pbinom(70, size = 73, prob = 0.93)
```


X < 0.9 x 70
X < 63
P(X < 63) = P(X <= 62) = pbinom(62, size = 73, prob = 0.93) = 0.01242371
```{r}
pbinom(62, size = 73, prob = 0.93)
```


Expected number of passengers that show up = E[X] = nπ = 73 x 0.93 = 67.89
```{r}
73 * 0.93
```



n = 73
π = 0.93
```{r}
# Expected Payment to Passengers = 200*1*P(X=71) + 200*2*P(X=72) + 200*3*P(X=73) = 28.89688
200*1*dbinom(71,73,0.93) + 200*2*dbinom(72,73,0.93) + 200*3*dbinom(73,73,0.93)
```



```{r}
# Let X be the binomial random variable denoting the number of passengers that show up.
# π = 0.93  i.e. Probability that a passenger shows up.
# As we increase the number of tickets sold -> n, say overbooking by m,
# The expected Revenue = 150 * (# of tickets) * P(X=70|n) + ∑ (150 * n - 200 * m)P(X=m+70) 
# summed over all m between 1 and m.
# i.e. either total passengers showing up make up 70 or less
# or # of passengers showing up equals any number between 71 and the overbooked number.
# I wrote the below commands to plot a graph and present in tabular format the revenue as we sell 71 to 90 tickets for the 70 seats.
# In the Graph and Table, we see that the Expected Revenue is maximum when we sell 77 tickets, i.e. overbook by 7.

ticketsSold <- seq(71, 90)
revenue <- vector(length = 20)
for (i in 1:20) {
  # Expected revenue when passengers that showed up total to 70
  revenue[i] = (150 * ticketsSold[i] * pbinom(70,ticketsSold[i],0.93))
  j <- i 
  # Calculating Expected Revenue for a high turn out for a given # of tickets sold
  revenueHighTurnOut = 0
    while (j > 0) {
      k = 70 + j
      revenueHighTurnOut = revenueHighTurnOut + (150 * ticketsSold[i] - 200 * j) * dbinom(k, ticketsSold[i], 0.93)
      j = j - 1
    }
  revenue[i] = revenue[i] + revenueHighTurnOut
  }
plot(ticketsSold, revenue, xlab = "# of tickets sold", ylab = "Expected Revenue", main = "Expected Revenue vs. Number of tickets sold")
data.frame(ticketsSold, revenue)
```


P(X>70 | X>= 70) = P( X > 70 AND X >= 70 ) / P(X >= 70)
P(X > 70) = (P( X <= 73 ) - P( X <= 70 )) / (1 - P(X <= 69)) = 0.446365

```{r}
(pbinom(73, 73, 0.93) - pbinom(70, 73, 0.93))/(1 - pbinom(69, 73, 0.93))
```



Business Class Seats = 4
Economy Class Seats = 66
Total Seats = 70
Business Class tickets sold = 6
Economy Class tickets sold = 68
P(Business Class passengers show up) = 0.75
P(Economy Class passengers show up) = 0.93
Let B be the number of Business Class passengers that show up
and E the number of Economy Class passengers that show up
P(Total_Passengers>72) = P(B=5)xP(E=68) + P(B=6)xP(E=67) + P(B=6)xP(E=68)
= dbinom(5, size = 6, 0.75)*dbinom(68, size = 68, 0.93) +
  dbinom(6, size = 6, 0.75)*dbinom(67, size = 68, 0.93) +
  dbinom(6, size = 6, 0.75)*dbinom(68, size = 68, 0.93)
= 0.01039135

```{r}
dbinom(5, size = 6, prob = 0.75)*dbinom(68, size = 68, prob = 0.93) +
  dbinom(6, size = 6, prob = 0.75)*dbinom(67, size = 68, prob = 0.93) +
  dbinom(6, size = 6, prob = 0.75)*dbinom(68, size = 68, prob = 0.93)
```

or we could also think of the problem as:
P(Total_Passengers>72) = P(B>4)xP(E>66) - P(B=5)xP(E=67) = 0.01039135
```{r}
(1 - pbinom(4, size = 6, prob = 0.75)) * (1 - pbinom(66, size = 68, prob = 0.93)) - (dbinom(5, size = 6, prob = 0.75) * dbinom(67, size = 68, prob = 0.93))
```

# Geometric Distribution


X = 0, 1, 2, .. , ∞


X = 0 when we roll a 3 on the first try.
Probability of getting a 3: P(3) = 1/6
Thus, P(X = 0) = P(3 on first try) = P(3) = 1/6 = 0.1666667


X = 1  i.e. when we didn't roll 3 on the first try but did on the second.
P(X=1) = (1 - P(3)) x P(3) = 5/6 x 1/6 = 5/36 = 0.1388889


P(X=2) = (1 - P(3)) x (1 - P(3)) x P(3) = 5/6 x 5/6 x 1/6 = 25/216 = 0.1157407



```{r}
# P(X = 0)
dgeom(0, prob = 1/6)

# P(X = 1)
dgeom(1, prob = 1/6)

# P(X = 2)
dgeom(2, prob = 1/6)
```


P(purchasing a computer) = 0.03
Let X be the variable denoting the number of unsuccessful trials.
Thus, the first success is after 10 unsuccessful trials or ater 10 customers who didn't buy.
We get the required probability by dgeom.
P(10) = dgeom(10, 0.03) = 0.02212272
```{r}
dgeom(10, prob = 0.03)
```



dgeom(30) = pgeom(30) - pgeom(29) = 0.01203021
```{r}
pgeom(30, prob =  0.03) - pgeom(29, prob =  0.03)
```

Yes, we can try to model using binomial distribution.
The order matters here and there can be only one order, unsuccessful trials followed by one successful buy, whereas in modelling using a binomial distribution, the order doesn't matter. 

To correct the binomial distribution model we need to divide its answer by the number of ways in which the events can be arranged when their order doesn't matter, i.e. nCr = 31C30 = 31

let X be the random number of customers that bought the computer.
Probability that a customer buys = π = 0.03
Number of trials = n = 31
Thus X is a Random Variable with n = 31 and π = 0.03
P(X=1|n, π) = dbinom(1, size = 31, prob = 0.03)/31 = 0.01203021
```{r}
dbinom(1, size = 31, prob = 0.03)/31
```


Let X be the discrete random variable denoting the number of customers that arrive before the customer that buys.
E[X] = (1 − π)/π = 32.33333

```{r}
# Expected number of customers that arrive before the successful event = mean of the random variable = (1 − π)/π
(1 - 0.03)/0.03
```
```{r}
# variance
(1 - 0.03)/(0.03)^2
```


Arithmetic Mean of 10000 rgeom numbers = 32.8398
Expected value using formula from h = 32.33333
We know that Arithmetic Mean of a progression > Geometric Mean
Thus, AM(32.5181) > GM(32.33)
The expected value calculation for Geometric Distribution has a geometric progression.
```{r}
rgeom_nos <- rgeom(10000, prob = 0.03)
mean(rgeom_nos)
(1 - 0.03)/0.03
```

