---
title: "Probability-Sampling-Distributions"
author: "Rahul Sinha"
date: 2018-10-03
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

# Poisson Distributions

Let N be the number of customers that arrive in the first 10 minutes that the store is open.
N follows a Poisson distribution, with lambda = 4

Question 1a asks for: P( N < 6 ) = P(N<=5)

```{r}
ppois(5,lambda = 4)
```

Let X be a random number denoting the additional number of scoops after the first scoop that a customer orders.

Given: X is binomial distributed with size(n) = 2, Probability(π) = 0.5
Let Y be the total number of scoops of ice cream that a randomly chosen customer will purchase.
We need to calculate the PMF for Y.
We know that Y takes values: 1, 2, 3
i.e. X takes values: 0, 1, 2
Y is always 1 + X

The PMF for X is given by: nCx (0.5)^x (0.5)^(n-x)
where x = 0, 1, 2
which comes to:
    PMF = 1/4 for x = 0,2
and PMF = 1/2 for x = 1


PMF for the number of scoops a customer will purchase:
X   Y   (PMF for X)   Expected number of scoops(1 + E[X])
0   1       1/4           1.25
1   2       1/2           1.50
2   3       1/4           1.25


Let X be the number of customers arriving at ice cream shop in 60 min
1 customer every 150 seconds --> 1/150 x 3600 = 24
X is a poisson distributed random variable with lambda = 24

```{r}
lambda_1hr <- 24
```
```{r}
sim_scoopsSold <- replicate( 10000, sum( rpois(1,lambda_1hr) * 0.25 + rpois(1,lambda_1hr) * 0.5 * 2 + rpois(1,lambda_1hr) * 0.25 *3 ) )
mean(sim_scoopsSold)
```


# Poisson Distributions
Let X be the number of cars arriving at the toll plaza in the first 10 minutes.

3 Cars at the toll plaza every 1 minute --> 30 Cars expected in the first 10 minutes.

```{r}
lambdaCar_10min <- 30
```


1.5 Trucks every 1 minute
```{r}
lambdaTruck_1min <- 1.5
dpois(0,lambdaTruck_1min)
```


expected value of number of vehicles in 20 min

3 Cars at the toll plaza every 1 minute --> 60 Cars expected in the first 20 minutes.
1.5 Trucks every 1 minute --> 30 Trucks expected in the first 20 minutes.
```{r}
lambdaCar_20min <- 60
lambdaTruck_20min <- 30
```

Let Z be the number of vehicles in first 20 mins


Let X be the number of Cars and Y be the number of Trucks in first 20 mins.

Z = X + Y

```{r}
sim_number_of_vehicles <- replicate( 10000 , rpois(1,lambdaCar_20min) + rpois(1,lambdaTruck_20min) )
```


expected number of vehicles:

```{r}
ev_vehicles <- mean(sim_number_of_vehicles)
```

standard deviation of vehicles:

```{r}
var_vehicles <- sum(  ( sim_number_of_vehicles - ev_vehicles )^2 ) / (10000-1)
sd_vehicles  <- sqrt(var_vehicles)
sd_vehicles
```



Toll charges:
Cars    $3
Trucks  $6.50

Find: Probability that more than $1000 in toll in next 1 hr.

3 Cars at the toll plaza every 1 minute --> 180 Cars expected in the first 60 minutes.
1.5 Trucks every 1 minute --> 90 Trucks expected in the first 60 minutes.
```{r}
lambdaCar_60min <- 180
lambdaTruck_60min <- 90
```

Let Z be the toll from vehicles in first 60 mins


Let X be the number of Cars and Y be the number of Trucks in first 60 mins.

Z = (3 x X) + (6.5 x Y)

Sampling for X and Y to get the total toll:

```{r}
sim_toll_vehicles <- replicate( 10000 , 3 * rpois(1,lambdaCar_60min) + 6.5 * rpois(1,lambdaTruck_60min) )
```

trying out the first sample:
```{r}
3 * rpois(1,lambdaCar_60min) + 6.5 * rpois(1,lambdaTruck_60min) 
```

```{r}
indicate_toll <- sim_toll_vehicles > 1000
```

```{r}
mean(indicate_toll)
```

Thus P(Z > 1000) = 0.9573



# Exponential Distribution


Let X be the exponential random variable denoting the lifetime of a light bulb.
µ = 5 years --> rate = 1/5
P(X > 7) = 1 - P(X <= 7) = 0.246597


```{r}
rate = 1/5
1 - pexp(7, rate)
```

P(X > 4) = 1 - P(X <= 4) = 0.449329

```{r}
rate = 1/5
1 - pexp(4, rate)
```

Find: P(X>10|X>6)

P(X>10|X>6) = P(X>10 and X>6)/P(X>6)
if X > 10 then X > 6  thus,
P(X>10|X>6) = P(X>10)/P(X>6) = 0.449329
```{r}
(1 - pexp(10,rate))/(1 - pexp(6,rate))
```

Find: P(X>14|X>10)

P(X>14|X>10) = P(X>14 and X>10)/P(X>10)
if X > 14 then X > 10  thus,
P(X>14|X>10) = P(X>14)/P(X>10) = 0.449329
```{r}
(1 - pexp(14,rate))/(1 - pexp(10,rate))
```

Yes, the probabilities for the above 3 questions are same. So, the probability that the bulb lasts for another 4 years given time t has passed is the same as the probability that the bulb lasted for 4 years to start with. This results from the formula and can be shown:
consider ß to represent lambda.
we know PDF for exponential distribution is: f(x) = ße^(-ßx)

P(X> 4+t)/P(X>4) = ße^(-ß(4+t)) / ße^(-ßt)
= e^(-4ß)

Thus, we see that the conditional probability for the exponential function doesn't depend on time t that has passed.


# Uniform Distribution
Let the length of the steel beam in inches be denoted by the uniform random variable X.
Minimum value of the Continuous Random Variable = 8.7 inches
Maximum value of the Continuous Random Variable = 9.1 inches
Thus, P(X = 8.5) = 0 as 8.5 inches is outside the range of values of the Continuous Random Variable.
Also, P(X = x) = 0 for any particular value of x, if X is a Continuous Random Variable.

```{r}
dunif(8.5, min = 8.7, max = 9.1)
```


P(X = x) = 0 for any particular value of x, if X is a Continuous Random Variable.
Therefore, P(X = 8.9) = 0

Let X be the Continuous Random Variable denoting the length of the steel beam in inches.
P(8.92 <= X <= 9.01) = P(X <= 9.01) - P(X <= 8.92)

```{r}
punif(9.01,min = 8.7,max = 9.1) - punif(8.92,min = 8.7,max = 9.1)
```

Let X be the Continuous Random Variable denoting the length of the steel beam in inches.
P(8.5 <= X <= 9.01) = P(X <= 9.01) - P(X <= 8.5)
```{r}
punif(9.01,min = 8.7,max = 9.1) - punif(8.5,min = 8.7,max = 9.1)
```

Let Y = the sum of lengths of 101 randomly chosen beams
P(Y<= 900)

Sampling truck loads:

```{r}
sum( runif(101,min=8.7,max=9.1) )
```

```{r}
sim_truckloads <- replicate( 10000, sum( runif(101,min=8.7,max=9.1) ) )
```
```{r}
ind_fit_on_truck <- sim_truckloads < 900
```

```{r}
mean(ind_fit_on_truck)
```

P(Y<= 900) = 0.8229

Thus Cruz Inc. will not be comfortable using Gopal LLC as P(Y<= 900) < 0.90
i.e. Cruz Inc. wants to be 90% certain that 101 beams can be shipped on a single truck, but the calculated probability suggests they can only be 82.29% certain using Gopal LLC.




